A Study of Institutional Spending on Open Access Publication Fees in Germany

### 2016-06-10 article.Rmd
minor corrections

The study of institutional spending on open access journal articles has been limited for several reasons. The first is that payment of these charges is fragmented across the budgets of grant agencies, research institutions, and libraries, or is covered by personal budgets. A comprehensive 2015 survey asked 9,645 authors from various disciplines how they financed publication fees, and it revealed that the majority of the respondents had access to research funding or institutional support to cover these charges. By contrast, 12 % paid publication fees individually [@soap]. These results are consistent with similar findings from other studies. Previous studies also found that funding sources exist primarily in higher-income countries, mainly to support research articles in the biological and physical sciences [@Solomon_2011]. Personal budgets, however, are likely used to cover lower publication fees [@Solomon_2011; @Bj_rk_2015].

@@

To address the problems of fragmented spending on publication fees and the lack of transparency about what is being paid, some European research funders and research-performing institutions have recently begun to disclose their expenditures for publication as open data. To the best of our knowledge, the first research funders to provide such data were the Wellcome Trust [@wellcome_apc_12] and the Austrian Science Fund FWF [@fwf_apc_13]. The not-for-profit company Jisc followed this example by collecting data from UK universities [@Lawson_data_2015]. Disclosed as publicly available spreadsheets, these data-sets self-report expenditures along with bibliographic information, including title, journal and publisher, and a persistent identifier for the publisher's version. Curatorial efforts focussed on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers to the full text including the Digital Object Identifier (DOI) [@Neylon_2014; @woodward_2014]. A preliminary version of Jisc's cost data was examined by @Pinfield_2015. Although the average spending on publication fees remained stable across universities, the authors found large price variations, as well as a varying number of articles supported by UK universities between 2007 and 2014, findings which confirm earlier studies that collected price information from journal websites [@Solomon_2012].

@@

Data collection also involved obtaining cost data from related open data initiatives. To compare self-reported spending on open access journal articles by German universities and research organisations with that of other initiatives, we consulted the openly available data-sets from the Austrian Science Fund (FWF) [@fwf_apc_14; @fwf_apc_15], Jisc [@jisc_14; @jisc_15] and the Wellcome Trust [@wellcome_apc_13_14; @wellcome_apc_14_15]. For analysis, we obtained the overall publication fee spending on both fully and hybrid open access journal articles. In the case of FWF, we gathered the cost information from the accompanying spending reports. We used the spreadsheet data to calculate Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average foreign exchange reference rates provided by the European Central Bank. Our comparison of the open data initiatives focussed on the last two years: 2014 and 2015. Because Wellcome Trust's spending was reported for the fiscal periods 2013-2014 and 2014-2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data. We excluded articles from the analysis with missing information about the cost or the journal type. In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to € 4.861.772 from the analysis because no publication type was given in the data-set.

@@

The fees for all of the articles amounted to € `r sum(my_apc$euro)`, including value-added tax; the average payment was € `r mean(my_apc$euro)` (median =  € `r median(my_apc$euro)`) . Figure 2 presents the distribution of institutional spending on publications. We observed that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100` %) of the publication fees were paid in accordance with the DFG price cap of € 2,000. Most payments for publications ranged from € 1,000 to € 1,250.

@@

Figure 3 presents institutional spending per article and year. Large price variations can be observed. Publication fees that were paid by German universities and research organisations ranged from € `r min(my_apc$euro)` to € `r max(my_apc$euro)` (SD = `r sd(my_apc$euro)`). However, the average price paid varied somewhat during the period from 2011 to 2015 (€ `r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - € `r my_apc %>% filter(period == 2015) %>% summarise(mean(euro)) %>% round()`).

@@

[^24]: <https://github.com/OpenAPC/openapc-de/tree/v2.4.3>

>>>>

The study of institutional spending on open access journal articles has been limited for several reasons. The first is that payment of these charges is fragmented across the budgets of grant agencies, research institutions, and libraries, or is covered by personal budgets. A comprehensive 2010 survey asked 9,645 authors from various disciplines how they financed publication fees, and it revealed that the majority of the respondents had access to research funding or institutional support to cover these charges. By contrast, 12 % paid publication fees individually [@soap]. These results are consistent with similar findings from other studies: Previous studies also found that funding sources exist primarily in higher-income countries, mainly to support research articles in the biological and physical sciences [@Solomon_2011]. Personal budgets, however, are likely used to cover lower publication fees [@Solomon_2011; @Bj_rk_2015].

@@

To address the problems of fragmented spending on publication fees and the lack of transparency about what is being paid, some European research funders and research-performing institutions have recently begun to disclose their expenditures for publication as open data. To the best of our knowledge, the first research funders to provide such data were the Wellcome Trust [@wellcome_apc_12] and the Austrian Science Fund FWF [@fwf_apc_13]. The not-for-profit company Jisc followed this example by collecting data from UK universities [@Lawson_data_2015]. Disclosed as publicly available spreadsheets, these data-sets self-report expenditures along with bibliographic information, including title, journal and publisher, and a persistent identifier for the publisher's version. Curatorial efforts focussed on the disambiguation of publisher and journal titles as well as on detecting duplicates and persistent identifiers for the full text including the Digital Object Identifier (DOI) [@Neylon_2014; @woodward_2014]. A preliminary version of Jisc's cost data was examined by @Pinfield_2015. Although the average spending on publication fees remained stable across universities, the authors found large price variations, as well as a varying number of articles supported by UK universities between 2007 and 2014, findings which confirm earlier studies that collected price information from journal websites [@Solomon_2012].

@@

Data collection also involved obtaining cost data from related open data initiatives. To compare self-reported spending on open access journal articles by German universities and research organisations with that of other initiatives, we consulted the openly available data-sets from the Austrian Science Fund (FWF) [@fwf_apc_14; @fwf_apc_15], Jisc [@jisc_14; @jisc_15] and the Wellcome Trust [@wellcome_apc_13_14; @wellcome_apc_14_15]. For analysis, we obtained the overall publication fee spending on both fully and hybrid open access journal articles. In the case of FWF, we gathered the cost information from the accompanying spending reports. We used the spreadsheet data to calculate Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average foreign exchange reference rates provided by the European Central Bank. Our comparison of the open data initiatives focussed on the last two years: 2014 and 2015. Because Wellcome Trust's spending was reported for the fiscal periods 2013-2014 and 2014-2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data. We excluded articles from the analysis with missing information about the cost or the journal type. In the case of Jisc's 2014 data [@jisc_14], for instance, we excluded spending on 2,812 publications that amounted to € 4,861,772 from the analysis because no publication type was given in the data-set.

@@

The fees for all of the articles amounted to € `r sum(my_apc$euro)`, including value-added tax; the average payment was € `r mean(my_apc$euro)` (median =  € `r median(my_apc$euro)`, SD = € `r sd(my_apc$euro)`). Figure 2 presents the distribution of institutional spending on publications. We observed that `r nrow(my_apc[my_apc$euro < 2000,])` (`r (nrow(my_apc[my_apc$euro < 2000,]) / nrow(my_apc)) *100` %) of the publication fees were paid in accordance with the DFG price cap of € 2,000. Most payments for publications ranged from € 1,000 to € 1,250.

@@

Figure 3 presents institutional spending per article and year. Large price variations can be observed. Publication fees that were paid by German universities and research organisations ranged from € `r min(my_apc$euro)` to € `r max(my_apc$euro)`. However, the average price paid varied somewhat during the period from 2011 to 2015 (€ `r my_apc %>% filter(period == 2011) %>% summarise(mean(euro)) %>% round()` - € `r my_apc %>% filter(period == 2015) %>% summarise(mean(euro)) %>% round()`).

@@

[^24]: <hhttps://github.com/OpenAPC/openapc-de/releases/tag/v2.4.3>
